---
title: "Untitled"
output: html_document
---
# Load Packages
```{r}
library(dplyr)
library(tidyr)
library(knitr)
library(ggplot2)
library(plotly)
```

# Read Data
```{r}
Non_Motorists <-read.csv("https://data.montgomerycountymd.gov/api/views/n7fk-dce5/rows.csv?accessType=DOWNLOAD")

Drivers <- read.csv('https://data.montgomerycountymd.gov/api/views/mmzv-x632/rows.csv?accessType=DOWNLOAD')

Incidents <- read.csv('https://data.montgomerycountymd.gov/api/views/bhju-22kf/rows.csv?accessType=DOWNLOAD')
```

# Group Municipality and crash.Date.Time, then count the number of drivers involved in incident by day (day/month/year) 
```{r}
Drivers$Crash.Date.Time<-substring(as.character(Drivers$Crash.Date.Time), 1, 10)
Drivers_1 <-Drivers %>% group_by(Municipality, Crash.Date.Time) %>% summarise(Drivers_Count = n())
```

# Group Municipality and crash.Date.Time, then count the number of incidents by day (day/month/year) 
```{r}
Incidents$Crash.Date.Time <-substring(as.character(Incidents$Crash.Date.Time), 1, 10)
Incidents_1 <-Incidents %>% group_by(Municipality, Crash.Date.Time) %>% summarise(Incidents_Count = n())
```

# Merge Drivers_1和Incidents_1
```{r}
Drivers_Incidents <-merge(Drivers_1, Incidents_1, by.x = c("Municipality", "Crash.Date.Time"), by.y = c("Municipality", "Crash.Date.Time"))
```

# Quantify injury severity: 5 means fatal injury, 4 means suspected serious, 3 means possible injury, 2 means suspected minor, 1 means no apparent injury 
```{r}
Non_Motorists$Injury.Severity <-recode(Non_Motorists$Injury.Severity, "FATAL INJURY" = 5, "SUSPECTED SERIOUS INJURY" = 4, "POSSIBLE INJURY" = 3, "SUSPECTED MINOR INJURY" = 2, "NO APPARENT INJURY" = 1)
```

# Quantify weather variable: 1 means good weather, 0 means bad weather. Remove NA, Unknown, other
```{r}
Non_Motorists$Weather <-
        recode(Non_Motorists$Weather, "BLOWING SNOW" = 0, "CLEAR" = 1, "CLOUDY" = 1, "FOGGY" = 0, "N/A" = -99, "OTHER" = -99, "RAINING" = 0, "SEVERE WINDS" = 0, "SLEET" = 0, "SNOW" = 0, "UNKNOWN" = -99, "WINTRY MIX" = 0)

Non_Motorists <- Non_Motorists[!Non_Motorists$Weather == -99, ]
```

# Quantify light condition, 1 means good visibility, 0 means bad visibility. Remove NA, Unknown, others
```{r}
Non_Motorists$Light <-recode(Non_Motorists$Light, "DARK -- UNKNOWN LIGHTING" = 0, "DARK LIGHTS ON" = 0, "DARK NO LIGHTS" = 0, "DAWN" = 0, "DAYLIGHT" = 1, "DUSK" = 0, "N/A" = -99, "OTHER" = -99, "UNKNOWN" = -99)

Non_Motorists <- Non_Motorists[!Non_Motorists$Light == -99, ]
```

# Quantify surface condition, 1 means good condition, 0 means bad condition Remove NA, Unknown, others
```{r}
Non_Motorists$Surface.Condition <-recode(Non_Motorists$Surface.Condition," " = -99,"DRY" = 1, "ICE" = 0, "N/A" = -99, "OTHER" = -99, "SLUSH" = 0, "SNOW" = 0, "UNKNOW" = -99, "WET" = 0)
Non_Motorists <-Non_Motorists[!Non_Motorists$Surface.Condition == -99, ]
```

# Quantify traffic condition, 1 means there is traffic control, 0 means there is no. Remove NA, Unknown, others
```{r}
Non_Motorists$Traffic.Control <-recode(Non_Motorists$Traffic.Control, "FLASHING TRAFFIC SIGNAL" = 1, "N/A" = -99, "NO CONTROLS" = 0, "OTHER" = -99, "PERSON" = 1, "RAILWAY CROSSING DEVICE" = 1, "SCHOOL ZONE SIGN DEVICE" = 1, "STOP SIGN" = 1, "TRAFFIC SIGNAL" = 1, "UNKNOWN" = -99, "WARNING SIGN" = 1, "YIELD SIGN" = 1)
Non_Motorists <- Non_Motorists[!Non_Motorists$Traffic.Control == -99, ]
```
# Quantify whether Non-Motorists is at fault, 1 at fault, 0 not. Remove NA, Unknown, others
```{r}
Non_Motorists$At.Fault <-recode(Non_Motorists$At.Fault,"No" = 0,"Unknown" = -99,"Yes" = 1)
Non_Motorists <- Non_Motorists[!Non_Motorists$At.Fault == -99, ]
```

# Turn Crash.Date.Time to month/day/year format
```{r}
Non_Motorists$Crash.Date.Time <-substring(as.character(Non_Motorists$Crash.Date.Time), 1, 10)
```

# Form new Non-Motorist dataset
```{r}
Non_Motorists_1 <-Non_Motorists %>% select(Municipality, Crash.Date.Time, Injury.Severity, Weather, Light, Surface.Condition, Traffic.Control, At.Fault)
```

# Merge Non_Motorists and Drivers_Incidents dataset and remove remaining NAs
```{r}
Non_Motorists_new <-
        merge(
                Non_Motorists_1,
                Drivers_Incidents,
                by.x = c("Municipality", "Crash.Date.Time"),
                by.y = c("Municipality", "Crash.Date.Time")
          
        )
```

```{r}
Data <- subset(Data,Municipality!="N/A")
```

```{r}
Data <- na.omit(Non_Motorists_new)
```

# View the final cleaned data
```{r}
View(Data)
```

# Add Daylight Saving Dates
```{r}
Data$daylight <- ifelse(
        (strptime(Data$Crash.Date.Time,'%m/%d/%Y')==strptime('2015-08-03','%Y-%m-%d') | strptime(Data$Crash.Date.Time,'%m/%d/%Y')==strptime('2015-11-01','%Y-%m-%d')) | (strptime(Data$Crash.Date.Time,'%m/%d/%Y')==strptime('2016-03-13','%Y-%m-%d') | strptime(Data$Crash.Date.Time,'%m/%d/%Y')==strptime('2016-11-06','%Y-%m-%d')) |
                (strptime(Data$Crash.Date.Time,'%m/%d/%Y')==strptime('2017-03-12','%Y-%m-%d') | strptime(Data$Crash.Date.Time,'%m/%d/%Y')==strptime('2017-11-05','%Y-%m-%d')) |
                (strptime(Data$Crash.Date.Time,'%m/%d/%Y')==strptime('2018-03-11','%Y-%m-%d') | strptime(Data$Crash.Date.Time,'%m/%d/%Y')==strptime('2018-11-04','%Y-%m-%d')) |
                (strptime(Data$Crash.Date.Time,'%m/%d/%Y')==strptime('2019-03-10','%Y-%m-%d') | strptime(Data$Crash.Date.Time,'%m/%d/%Y')==strptime('2019-11-03','%Y-%m-%d')) |
                (strptime(Data$Crash.Date.Time,'%m/%d/%Y')==strptime('2020-03-08','%Y-%m-%d') | strptime(Data$Crash.Date.Time,'%m/%d/%Y')==strptime('2020-11-01','%Y-%m-%d')) |
                (strptime(Data$Crash.Date.Time,'%m/%d/%Y')==strptime('2021-03-14','%Y-%m-%d') | strptime(Data$Crash.Date.Time,'%m/%d/%Y')==strptime('2021-11-07','%Y-%m-%d')),1,0
)
```

# Exploratory Data Analysis
# The new dataset contains 10 variables，Injury.Severity, Weather, Light, Surface.Condition, Traffic.Control, At.Fault are categorical variables
# Drivers_Count, Incidents_Count are quantitative
# Frequency counts for categorical variables

# Frequency count for Weather，0 represent bad weather，1 means good weather
```{r}
Data %>% 
        count(Weather) %>% 
        mutate(prop = prop.table(n))
```

# Frequency count for Light，1 represent good visibility，0 means bad visibility
```{r}
Data %>% 
        count(Light) %>% 
        mutate(prop = prop.table(n))
```
# Frequency count Surface.Condition, 1 means normal condition，0 means abnormal condition
```{r}
Data %>% 
        count(Surface.Condition) %>% 
        mutate(prop = prop.table(n))
```
# Frequency counts for Traffic Control, 1 means there is control，0 mean no
```{r}
Data %>% 
        count(Traffic.Control) %>% 
        mutate(prop = prop.table(n))
```
# Frequency count for wheather Non-Motortist is at fault, 1 means at fault，0 means not
```{r}
Data %>% 
        count(At.Fault) %>% 
        mutate(prop = prop.table(n))
```
# Frequency counts for injury severity
```{r}
Data %>% 
        count(Injury.Severity) %>% 
        mutate(prop = prop.table(n))
```
# Contingency table for the variables and injury severity, trying to find relationship

# Weather and Injury severity
```{r}
Data %>% 
        group_by(Injury.Severity, Weather) %>% 
        summarize(n = n()) %>% 
        spread(Weather, n) %>%
        kable()
```

# Light and Injury severity
```{r}
Data %>% 
        group_by(Injury.Severity, Light) %>% 
        summarize(n = n()) %>% 
        spread(Light, n) %>%
        kable()
```

# Surface condition and Injury severity
```{r}
Data %>% 
        group_by(Injury.Severity, Surface.Condition) %>% 
        summarize(n = n()) %>% 
        spread(Surface.Condition, n) %>%
        kable()
```

# Traffic control and Injury severity
```{r}
Data %>% 
        group_by(Injury.Severity, Traffic.Control) %>% 
        summarize(n = n()) %>% 
        spread(Traffic.Control, n) %>%
        kable()
```

# At fault and Injury severity
```{r}
Data %>% 
        group_by(Injury.Severity, At.Fault) %>% 
        summarize(n = n()) %>% 
        spread(At.Fault, n) %>%
        kable()
```

# Max, Min, Average, Standard Deviation for the quantitative variables

```{r}
Data %>% 
        select(Drivers_Count, Incidents_Count) %>%
        summarise(n = n(), Drivers_Count_average = mean(Drivers_Count), Incidents_Count_average = mean(Incidents_Count),
                  Drivers_Count_max = max(Drivers_Count), Incidents_Count_max = max(Incidents_Count),
                  Drivers_Count_min = min(Drivers_Count), Incidents_Count_min = min(Incidents_Count),
                  Drivers_Count_sd = sd(Drivers_Count), Incidents_Count_sd = sd(Incidents_Count)) 
```

# Boxplot for injury severity and incident count by day
```{r}
ggplot(Data, aes(x = factor(Injury.Severity, labels = c("NO ","MINOR ","POSSIBLE", "SERIOUS","FATAL")), y = Incidents_Count)) +
        geom_boxplot(varwidth=TRUE, fill="plum") +
        labs(title="Distribution of Incidents Count For Days by Injury Severity",
             caption="data.montgomerycountymd.gov",
             x="Injury Severity",
             y="Incidents Count")
```
# Visibility's effect on Injury.Severity and incident count
```{r}
Data$Visibility <- factor(Data$Light,labels=c("Bad","Good"))
ggplot(Data, aes(x = factor(Injury.Severity, labels = c("NO","MINOR","POSSIBLE",
        "SERIOUS","FATAL")), y = Incidents_Count, fill=Visibility )) +
        geom_boxplot(varwidth=TRUE) +
        labs(title="Light Condition's Conbribution To Injury Severity and Count",
             caption="data.montgomerycountymd.gov",
             x="Injury Severity",
             y="Incidents Count")
```
# Weather's effect on Injury.Severity and incident count
```{r}
Data$Climate <- factor(Data$Weather,labels=c("Bad Weather","Good Weather"))
ggplot(Data, aes(x = factor(Injury.Severity, labels = c("NO","MINOR","POSSIBLE","SERIOUS","FATAL")), y = Incidents_Count,
                 fill=Climate )) +
        geom_boxplot(varwidth=TRUE) +
        labs(title="Weather's Conbribution To Injury Severity and Count",
             caption="data.montgomerycountymd.gov",
             x="Injury Severity",
             y="Incidents Count")
```
# At Fault's Conbribution To Injury Severity and Count
```{r}
Data$Fault <- factor(Data$At.Fault,labels=c("Yes","No"))
ggplot(Data, aes(x = factor(Injury.Severity, labels = c("NO","MINOR","POSSIBLE",
        "SERIOUS","FATAL")), y = Incidents_Count, fill=Fault )) +
        geom_boxplot(varwidth=TRUE) +
        labs(title="At Fault's Conbribution To Injury Severity and Count",
             caption="data.montgomerycountymd.gov",
             x="Injury Severity",
             y="Incidents Count")
```
# Trend Analysis of the Number of Injuries, Month, and Municipality
```{r}
s1 <- Data%>%group_by(Municipality,Injury.Severity, strftime(strptime(Crash.Date.Time,'%m/%d/%Y'),'%m'))%>%summarise(n=n())
colnames(s1) = c("Municipality", "Severity","time", "n")

ggplot(s1, aes(time, n)) +
        geom_point(aes(colour = Municipality)) +
        facet_grid(factor(Severity, labels = c("NO","MINOR ","POSSIBLE", "SERIOUS","FATAL"))~.) +
        labs(title = "Trend Analysis of the Number of Injuries, Month, and Municipality",
             caption="data.montgomerycountymd.gov",
             x="Month", y ="Non-motorist injuries")
```

# Number of injuries, Year, Municipality, and At Fault
```{r}
s2 <- Data%>%group_by(Municipality,Injury.Severity,At.Fault, strftime(strptime(Crash.Date.Time,'%m/%d/%Y'),'%Y'))%>%summarise(n=n())
colnames(s2) = c("Municipality", "Severity","Surface","time", "n")

ggplot(s2, aes(time, n))+
        geom_point(aes(colour = Municipality)) +
        facet_grid(factor(Surface, labels = c('YES','NO'))~factor(Severity, labels = c("NO","MINOR ","POSSIBLE ",
                                                                                         "SERIOUS ","FATAL"))) +
        labs(title = "Number of injuries, Year, Municipality, and At Fault",
             caption="data.montgomerycountymd.gov",
             x="Year", y ="Non-motorised injuries")
```

# Statistical Analysis

# Using Chi-squared test，evaluate Injury Severity and categorical variables's 
```{r}
chisq.test(Data$Injury.Severity,Data$Weather)
chisq.test(Data$Injury.Severity,Data$Light)
chisq.test(Data$Injury.Severity,Data$Surface.Condition)
chisq.test(Data$Injury.Severity,Data$Traffic.Control)
chisq.test(Data$Injury.Severity,Data$At.Fault)
```
# Analysis: The p-vaule for light condition, traffic control, and at fault are smaller than 0.05, which mean they might be statistically significantly associated with injury severity.


# Use Shapiro-Wilk normality test if Drivers and Incidents count are normally distributed
```{r}
shapiro.test(Data$Drivers_Count)
shapiro.test(Data$Incidents_Count)
```
# Result: Both W-scores are close to 0 but p-value is less than 0.5. So they are not normally distributed. 


# Multiple linear regression
```{r}
lm.mol <- lm(Injury.Severity~Weather+Light+Surface.Condition+Traffic.Control+At.Fault, data =Data)
summary(lm.mol)
```
# By making injury severity as a dependent variable, the resulting R^2 value of both multiple and adjusted are between 0 and 1 (0.1162 and 0.0973), meaning there is a relatively strong correlation between injury severity and the variables. The p-value is less than 0.05 being 1.303e-05, it is statistically significant. The variables Light, At fault, and daylight are showing statistical significance.










































